---
title: "Lab 2"
author: "Your Name Here"
date: ""
output:
  html_document:
    toc: true
    toc_depth: 4
    theme: cerulean
    highlight: tango
---

```{r package_load}
library(ggplot2) # graphics library
library(MASS)    # contains data sets
library(ISLR)    # contains code and data from the textbook
library(knitr)   # contains kable() function
library(boot)    # contains cross-validation functions
library(gam)     # needed for additive models

options(scipen = 4)  # Suppresses scientific notation
```

### 1. Changing the author field and file name.

##### (a) Change the `author:` field on the Rmd document from Your Name Here to your own name.

##### (b) Rename this file to "lab01_YourHameHere.Rmd", where YourNameHere is changed to your own name.

<hr>

> The next portion of the lab gets you to carry out the Lab in &sect;5.3 of ISLR (Pages 191 - 193).  You will want to have the textbook Lab open in front you as you go through these exercises.  The ISLR Lab provides much more context and explanation for what you're doing.  



### 2. The Validtion Set Approach

> You will need the `Auto` data set from the `ISLR` library in order to complete this exercise.

> Please run all of the code indicated in &sect;5.3.1 of ISLR, even if I don't explicitly ask you to do so in this document.

##### (a) Run the `View()` command on the `Auto` data to see what the data set looks like.

```{r}
#View(Auto)
```

##### Use `qplot` to construct a scatterplot of `mpg` vs `horsepower`.  Use `stat_smooth()` to overlay a linear, quadratic, and cubic polynomial fit to the data.  

```{r}
qplot(data = Auto, x = horsepower, y = mpg) +
  stat_smooth(method = "lm", aes(colour = "Linear regression")) + 
  stat_smooth(method = "lm", formula = y ~ poly(x, 2), aes(colour = "Quadratic fit")) +
  scale_colour_discrete("Model") +
  theme_bw()
```

##### (b) Use the command `set.seed(1)` to set the seed of the random number generator.  This will ensure that your answers match those in the text.

```{r}
# Edit me
set.seed(1)
```

##### (c) Use the `sample()` command to construct `train`, a vector of observation indexes to be used for the purpose of training your model.

```{r}
# Edit me
train <- sample(392, 196)
```

##### Describe what the sample() function as used above actually does.

- When used in `sample(n, size)` syntax, the `sample` function produces a random sample of size `size` from `1:n`.  Sampling is done *without replacement*.  

##### (c) Fit a linear model regression `mpg` on `horsepower`, specifying `subset = train`.  Save this in a variable called `lm.fit`.

```{r}
# Edit me
lm.fit <- lm(mpg ~ horsepower, data = Auto, subset = train) 
```

##### Why do we use the argument `subset = train`?

- The `train` indexes form the training set.  We want to train on just the `train` data, not the full data.

##### (d) Calculate the MSE of `lm.fit` on the test set (i.e., all points that are not in `train`)

```{r}
mse1 <- with(Auto, mean((mpg - predict(lm.fit, Auto))[-train]^2))
mse1
```

##### (e) Use the `poly()` command to fit a quadratic regression model of `mpg` on `horsepower`, specifying `subset = train`.  Save this in a variable called `lm.fit2`.    

```{r}
lm.fit2 <- lm(mpg ~ poly(horsepower, 2), data = Auto, subset = train)
```

##### Is the coefficient of the quadratic term statistically significant?  

```{r}
# Pull the p-value
coef(summary(lm.fit2))["poly(horsepower, 2)2", "Pr(>|t|)"]
```

- Yes, the coefficient is highly statistically significant

##### Calculate the test MSE of `lm.fit2`.  How does it compare to the linear regression fit?

```{r}
mse2 <- with(Auto, mean((mpg - predict(lm.fit2, Auto))[-train]^2))
mse2
```

- The test error of the quadratic model is `r round(mse2, 1)`, which is considerably lower than the test error of the linear model `r round(mse1, 1)`.

##### (f) Use the `poly()` command to fit a cubic regression model of `mpg` on `horsepower`, specifying `subset = train`.  Save this in a variable called `lm.fit3`.    

```{r}
lm.fit3 <- lm(mpg ~ poly(horsepower, 3), data = Auto, subset = train)
```

##### Is the coefficient of the cubic term statistically significant?  

```{r}
# Pull the p-value
coef(summary(lm.fit3))["poly(horsepower, 3)3", "Pr(>|t|)"]
```

- No, the coefficient of the cubic term is not statistically significant.  It looks like going from quadratic to cubic doesn't help to explain much additional variability in the data

##### Calculate the test error of the cubic fit.

```{r}
mse3 <- with(Auto, mean((mpg - predict(lm.fit3, Auto))[-train]^2))
mse3
```

##### (g) How do the test errors of the three models compare?  Which of the three models should we use?  

```{r}
mse1 # Linear test error
mse2 # Quadratic test error
mse3 # Cubic test error
```

- While the cubic model has ever-so-slightly lower test error, it's very close to the test error of the quadratic model.  Given that the cubic term isn't even statistically significant, the **quadratic model** is the best model to use.  The quadratic does *much* better than the linear model.

##### (h) Re-run all of the code above, but this time setting `set.seed(5)`.  You do not have to retype all of the code.  You can just change the initial `set.seed()` command and see what happens.  

##### What does changing the seed value do?  Did this change your estimated test errors?  Why did the values change?  Should we still pick the same model?

- When you re-run the code with `set.seed(5)`, you are getting a different random split of the data between Train and Validation.  The estimates of test error also change, because we now have slightly different model estimates, and a different random set to test the models on.  We now get an error of `22.2` for the linear model; `15.2` for the quadratic, and `15.2` for the cubic.  These numbers are quite different from our estimates based on the first split we tried.  However, the conclusion is essentially the same:  The cubic model doesn't seem to add much, but the quadratic *is* much better than the linear.  We should use the quadratic model.

### 3. Leave-One-Out Cross-Validation

> This exercise introduces you to the `cv.glm()` command from the `boot` library, which automates K-fold cross-validation for Generalized Linear Models (GLMs).  Linear regression is one example of a GLM.  Logistic regression is another.  GLMs are not the same as Generalized Additive Models (GAMs).

> Please run all of the code indicated in &sect;5.3.2 of ISLR, even if I don't explicitly ask you to do so in this document.

##### (a) Use the `glm` command to fit a linear regression of `mpg` on `horsepower`.  Call the resulting model `glm.fit`  Confirm that this gives the same coefficient estimates as a linear model fit with the `lm` command.  

**Note**: You should fit the model to the entire data, not just to the training data.

```{r}
glm.fit <- glm(mpg ~ horsepower, data = Auto)

coef(glm.fit)
coef(lm(mpg ~ horsepower, data = Auto))

# Are they the same? Yes.
identical(coef(glm.fit), coef(lm(mpg ~ horsepower, data = Auto)))
```

##### (b) Run all of the code in &sect;5.3.2.   Construct a plot of `cv.error`, the vector of LOOCV error estimates for polynomials of degree 1-5.  

**Note**: The computations take some time to run, so I've set `cache = TRUE` in the code chunk header to make sure that the code below doesn't re-execute at knit time unless this particular chunk has changed.

```{r, cache = TRUE}
cv.err <- cv.glm(Auto, glm.fit)
cv.err$delta
cv.error <- rep(0, 5)
for (i in 1:5) {
  glm.fit <- glm(mpg ~ poly(horsepower, i), data=Auto)
  cv.error[i] <- cv.glm(Auto, glm.fit)$delta[1]
}
cv.error
```

```{r}
# Form a little data frame for plotting
cv.df <- data.frame(degree = 1:5,
                    cv.error = cv.error)

qplot(data = cv.df, x = degree, y = cv.error, geom = "line",
      ylab = "LOOCV error estimate") + geom_point()
```

##### (c) Which degree model has the lowest LOOCV estimate of test error?  

```{r}
which.min(cv.error)
```

- The degree `which.min(cv.error)` model has the lowest LOOCV estimate of test error.

##### (d) Which model should we choose?  Is this the model with the lowest LOOCV estimate of test error?  Explain.

- We should choose the quadratic model.  While this model doesn't have the lowest LOOCV estimate of test error, it's estimated test error is quite close to that of the degree-5 polynomial model.  Since there estimated test error of the quadratic and degree-5 model are quite close, we should go with the quadratic because it's the simpler model.

### 4. K-fold Cross-validation  

> Please run all of the code indicated in &sect;5.3.3 of ISLR

##### (a)  Run all of the code in the $k$-Fold Cross-validation Lab section.

```{r, cache = TRUE}
set.seed(17)
cv.error.10 <- rep(0,10)
for (i in 1:10){
  glm.fit=glm(mpg ~ poly(horsepower, i), data=Auto)
  cv.error.10[i] <- cv.glm(Auto, glm.fit, K=10)$delta[1]
}
cv.error.10
```

##### (b) Construct a plot of 10-fold CV error vs. degree.  

```{r}
# Form a little data frame for plotting
cv.df <- data.frame(degree = 1:10,
                    cv.error = cv.error.10)

qplot(data = cv.df, x = degree, y = cv.error, geom = "line",
      ylab = "10-fold CV error estimate") + geom_point()
```

##### (c) Which model has the lowest 10-fold CV estimate of test error?

```{r}
which.min(cv.error.10)
```

- The model with degree `r which.min(cv.error.10)` has the lowest 10-fold CV estimate of test error. 

##### (d) Which model should we choose?  Is this the model with the lowest 10-fold CV estimate of test error?  Explain.

- We should choose the quadratic model.  While this model doesn't have the lowest 10-fold CV estimate of test error, it's estimated test error is quite close to that of the degree-5 polynomial model.  Since there estimated test error of the quadratic and degree-5 model are quite close, we should go with the quadratic because it's the simpler model.

### 5. Additive Models and Splines

> Please run all of the code indicated in &sect;7.8.3 of ISLR, up until the loading of the `akima` package.  We have not yet studied logistic regression, so you are not asked to do the logistic regression analysis that starts at the bottom of p. 297.  The material on ANOVA testing may also be unfamiliar to you.  You may skip it.

##### (a) Think carefully about what each line of code below is doing.  Write comments in the code below to explain what each line is responsible for.  

```{r, fig.width = 8, fig.height = 4}
# Use the lm function to fit an additive model 
# (using df = 4 and df = 5 natural cubic splines)
gam1 <- lm(wage ~ ns(year, 4) + ns(age, 5) + education, data=Wage)

# Use gam function to fit addition model with smoothing splines (df = 4, 5)
gam.m3 <- gam(wage ~ s(year, 4) + s(age, 5) + education, data=Wage)

# Construct a figure with 1 row, 3 columns
par(mfrow=c(1,3))

# Use the plot.gam function to get plots of the estimated f_j
# Note: plot() on a gam object is an alias for plot.gam()
plot(gam.m3, se=TRUE, col="blue")

# See above.
plot.gam(gam1, se=TRUE, col="red")

# Consider model where year doesn't appear
gam.m1 <- gam(wage ~ s(age,5) + education, data=Wage)

# Consider model wehre year appears linearly
gam.m2 <- gam(wage ~ year + s(age, 5) + education, data=Wage)

# Run an analysis of variance (ANOVA) test to see if 
# the m2 is statistically significantly better than m1,
# and if m3 is statistically significant better than m2
# Note: m3, m2 and m1 are nested models
anova(gam.m1, gam.m2, gam.m3, test="F")

# Print-out summary of additive model
summary(gam.m3)

# Use gam.m2 to get fitted values
preds <- predict(gam.m2, newdata=Wage)

# Try local regression for age instead of smoothing spline
gam.lo <- gam(wage ~ s(year, df=4) + lo(age, span=0.7) + education, data=Wage)

# Show plots of estimated fits
plot.gam(gam.lo, se=TRUE, col="green")

# Interaction model, where local regression is used to fit the joint effect of
# (year, age)
gam.lo.i <- gam(wage ~ lo(year, age, span=0.5) + education, data=Wage)
```

##### (a) Looking at the fit plots, does it look like an additive model is a good fit for the data, or are we "overcomplicating things", and a linear model would've probably done fine also?

```{r}
par(mfrow=c(1,3))

# Use the plot.gam function to get plots of the estimated f_j
# Note: plot() on a gam object is an alias for plot.gam()
plot(gam.m3, se=TRUE, col="blue")
```

- Looking at hte standard error bands, the `age` plot clearly indicates that the fit is not consistent with being linear.  The `year` plot, however, shows error bands that easily accomodate a linear fit between the response and `year`.  Thus we should use a non-linear model for `age`, but would be well-served by a linear model for `year`.  

##### (b)  **Optional**:  Use the ANOVA test output described in the ISLR discussion of the lab to help better answer part (a).

```{r}
anova(gam.m1, gam.m2, gam.m3, test="F")
```

- The p-value for going from the `~ year + ...` model to the `~ s(4, year) + ...` model is not statistically significant.  The smoothing spline fit doesn't add sufficient additional explanatory power to warrant the increase in model complexity.  This agrees with our observation from part (a).

### 6. Splines and cross-validation

> The `splines` library has a `smooth.spline()` command with built-in cross-validated smoothness selection.  We will now give an example of using this command.

##### The code below is adapted from the bottom half of page 293.  Add comments to the code below indicating what each line of code is doing.

```{r}
agelims <- range(Wage$age)

# Start a scatterplot, to be overlaid with smoothing spline fits
par(mfrow = c(1,1))
with(Wage, plot(age, wage, xlim = agelims, cex=0.5, col = "darkgrey"))
title("Smoothing Spline")

# Fit smoothing spline model with 16 effective degrees of freedom
fit <- with(Wage, smooth.spline(age, wage, df=16))

# Fit smoothing spline, using LOOCV to figure out the optimal choice of df
fit2 <- with(Wage, smooth.spline(age, wage, cv=TRUE))

# CV error minimizing choice of df
fit2$df

# Overlay df = 16 and CV error minimizing df smoothing spline fits
lines(fit, col="red", lwd=2)
lines(fit2, col="blue", lwd=2)
```

##### Based on the documentation for the `smooth.spline` function, can you figure out what kind of cross-validation is done when `cv = TRUE`?  i.e., It's $K$-fold CV for what choice of $K$?

- From the documentation for `smooth.spline`: 

`leave-one-out cross-validation (ordinary or ‘generalized’ as determined by cv) is used`